
##title: "CaseStudy1
### DJ's Contribution

#### Installing  dplyr and ggplot functions


```{r}
install.packages("dplyr",repos = "http://cran.us.r-project.org")
install.packages("ggplot2",repos = "http://cran.us.r-project.org")

```
### Reading the Beer Data from Repo created by 

```{r}
URL <- "https://github.com/selwyn-samuel-smu/CaseStudy01/"
temp <- tempfile()
download.file(URL,temp)

```
### Reading the Breweries.csv file from the repo

```{r}
brewer <- read.csv("Breweries.csv",header = TRUE)  # read csv file
head(brewer)

```

### Check the dim,names and str

```{r}
dim(brewer)
names(brewer)
str(brewer)

```

### Reading the Beers.csv file from the repo

```{r}
beers <- read.csv("Beers.csv",header = TRUE)  # read csv file
head(beers)

```
### Checkinh the dim,names and str

```{r}
dim(beers)
names(beers)
str(beers)

```
```{r}
brewer
```
### Checking the data structure of column name and state
```{r}
is.factor(brewer$State)
is.factor(brewer$Name)
```
### Converting the state to a factor and 
### Creating a data frame brewer.Name and brewer.state
```{r}
State = as.factor(brewer$State)
Data1 <- data.frame(brewer$Name, brewer$State)
Data1
```

###Question: 1 How many breweries are present in each state?

```{r}
library(dplyr)
Data1 %>%  group_by(brewer.State) %>% summarise(brewer.Name_count =n())

```


### Question 2 Merging Data Sets

```{r}
beer_data <- merge(brewer,beers, by = "Brewery_id","Brew_ID")
beer_data
head(beer_data)

```

##Q3 - Number of NA's in each colunm

```{r}
 sapply(beer_data,function(x) sum(is.na(x)))

```


## question 4 Part 1 -ompute the median alchohol content and international bitterness unit of each srate. Plot a bar chart to compare
```{r}
Data3 <- data.frame(beer_data$State, beer_data$ABV,beer_data$IBU)
library(dplyr)
Data3 %>%  group_by(beer_data.State) %>% summarise(median(beer_data.ABV),median(beer_data.IBU))
Data3
```
### Removing the NA from the data set

```{r}

Data3 <- na.omit(Data3)
Data3

```


```{r}
str(Data4)

```






